---
title: "Analyzing Focus Group Conversations with the FocusGroup Package"
author: "Ali Sanaei"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Focus Group Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(FocusGroup)
```

# Introduction

The FocusGroup package provides tools for simulating and analyzing focus group conversations. This vignette demonstrates how to use the package to create, run, and analyze focus group simulations. The package uses `FGAgent`s as the agents who participate in a focus group study; these agents can be based on various LLM models potentially from different providers; see `LLMR` documentation for which providers have built-in support and how to add new ones. One `FGAgent` is always the moderator. 

## Installation

```{r install, eval = FALSE}
# Install from GitHub
remotes::install_github("asanaei/FocusGroup")
```

# Running a Simple Focus Group

Let's start by running a focus group with 5 participants using the high-level wrapper. Note that you need to set your OpenAI API key in the environment variable `OPENAI_API_KEY`:

```{r create_fg, eval = FALSE}
# Set your OpenAI API key (or set it in your environment)
Sys.setenv(OPENAI_API_KEY = "your-api-key-here")

# Run a focus group with 5 participants
result <- run_focus_group(
  topic = "Impact of Social Media on Mental Health",
  participants = 5
)

# View the focus group configuration
print(result$focus_group)
```

# Viewing the Conversation

```{r view_conversation, eval = FALSE}
# View the conversation
head(result$conversation)
```

# Analyzing the Results

The package provides several analysis functions to help understand the conversation:

## Basic Analysis

```{r basic_analysis, eval = FALSE}
# Get basic statistics
stats <- analyze_focus_group(result)
print(stats$summary)
```

## Text Analysis

```{r text_analysis, eval = FALSE}
# Analyze text complexity
complexity <- analyze_text_complexity(result)
print(complexity$readability)

# Analyze response patterns
patterns <- analyze_response_patterns(result)
print(patterns$turn_patterns)
```

## Question Analysis

```{r questions, eval = FALSE}
# Analyze question patterns
questions <- analyze_question_patterns(result)
print(questions$question_patterns)
```

## Visualization

The package includes several plotting functions:

```{r plots, eval = FALSE}
# Create participation timeline
participation_plot <- create_participation_timeline(result)
print(participation_plot)

# Create sentiment timeline
sentiment_plot <- create_sentiment_timeline(result)
print(sentiment_plot)

# Create readability by phase
readability_plot <- create_readability_by_phase(result)
print(readability_plot)
```

# Advanced Usage

## Custom Agent Profiles and Manual Construction

You can create custom agent profiles and manually construct a FocusGroup object:

```{r custom_agents, eval = FALSE}
# Set your OpenAI API key (or set it in your environment)
Sys.setenv(OPENAI_API_KEY = "your-api-key-here")

# Create diverse agents
agents <- create_diverse_agents(
  n_participants = 3,
  demographics = list(
    age = c(25, 35, 45),
    gender = c("male", "female", "non-binary"),
    education = c("high school", "bachelor's", "master's")
  )
)

# Name the agents by their IDs
agents_named <- setNames(agents, sapply(agents, function(a) a$id))

# Find moderator ID
moderator_id <- agents_named[[1]]$id

# Create a conversation flow object
flow_obj <- DesireBasedFlow$new(agents_named, moderator_id)

# Create the FocusGroup object
fg <- FocusGroup$new(
  topic = "Impact of Social Media on Mental Health",
  purpose = "To explore perspectives and experiences related to social media.",
  agents = agents_named,
  moderator_id = moderator_id,
  turn_taking_flow = flow_obj
)
```

## Custom Analysis

You can perform custom analysis using the conversation data:

```{r custom_analysis, eval = FALSE}
# Extract conversation data
conv_data <- result$conversation

# Custom analysis example
word_counts <- sapply(conv_data$message, function(x) length(strsplit(x, "\\s+")[[1]]))
summary(word_counts)
```

# Best Practices

1. Always start with a clear research question
2. Use appropriate number of participants (typically 5-8)
3. Set realistic duration for the simulation
4. Review and adjust agent profiles based on your research needs
5. Use multiple analysis methods to get a comprehensive understanding
6. Remember to set your API key in the environment variable like `OPENAI_API_KEY`

# Conclusion

The FocusGroup package provides a powerful toolkit for simulating and analyzing focus group conversations. By following this vignette, you should be able to:

- Create and configure focus groups
- Run simulations with different parameters
- Analyze conversations using various methods
- Visualize results effectively
- Customize the analysis to your specific needs

For more information, please refer to the package documentation and examples. 
