---
title: "Focus Group Analysis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Focus Group Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(FocusGroup)
```

# Analyze a simulated discussion

FocusGroup stores a moderated discussion in a `FocusGroup` object. Each
participant and the moderator is an `FGAgent`. The group object contains the
question script, conversation flow, messages, response metadata, and analysis
methods.

Provider calls use LLMR. Set the provider key in an environment variable before
running a session. LLMR returns response metadata and token counts, which
FocusGroup records with each generated message.

## Installation

```{r install, eval = FALSE}
# Install from GitHub
remotes::install_github("asanaei/FocusGroup")
```

# Run a focus group

`run_focus_group()` constructs five participants, a moderator, a script, and a
conversation flow in the following example.

```{r create_fg, eval = FALSE}
# Set your OpenAI API key (or set it in your environment)
Sys.setenv(OPENAI_API_KEY = "your-api-key-here")

# Run a focus group with 5 participants
result <- run_focus_group(
  topic = "Impact of Social Media on Mental Health",
  participants = 5
)

# View the focus group configuration
print(result$focus_group)
```

# Read the conversation log

`result$conversation` is a data frame with one row per stored message. It
includes speaker and phase identifiers together with provider metadata and
token counts.

```{r view_conversation, eval = FALSE}
# View the conversation
head(result$conversation)
```

# Analyze the transcript

`analyze_focus_group()` runs the main analyses and returns them in a list. Topic,
readability, and plotting components require their suggested packages.
Thematic analysis makes a model call.

## Basic analysis

```{r basic_analysis, eval = FALSE}
# Get basic statistics
stats <- analyze_focus_group(result)
print(stats$basic_stats)
```

## Text analysis

Individual analyses are methods on `result$focus_group`.

```{r text_analysis, eval = FALSE}
# Readability scores per speaker
readability <- result$focus_group$analyze_readability()
print(readability)

# Response patterns and interaction behaviors
patterns <- result$focus_group$analyze_response_patterns()
print(patterns$response_metrics)
```

## Question analysis

```{r questions, eval = FALSE}
# Analyze question patterns
questions <- result$focus_group$analyze_question_patterns()
print(questions$question_patterns)
```

## Plots

Plotting methods return ggplot objects. Analysis methods continue to return
tables and lists.

```{r plots, eval = FALSE}
# Participation timeline
participation_plot <- result$focus_group$plot_participation_timeline()
print(participation_plot)

# Readability by speaker (a table, not a plot)
print(result$focus_group$analyze_readability())
```

# Construct the objects directly

The R6 interface exposes the same parts assembled by `run_focus_group()`. Create
the agents, name the list by agent identifier, construct a conversation flow,
and pass both objects to `FocusGroup$new()`.

```{r custom_agents, eval = FALSE}
# Set your OpenAI API key (or set it in your environment)
Sys.setenv(OPENAI_API_KEY = "your-api-key-here")

# Create diverse agents. `demographics` is a data frame with one row per
# participant.
agents <- create_diverse_agents(
  n_participants = 3,
  demographics = data.frame(
    age = c(25, 35, 45),
    gender = c("male", "female", "non-binary"),
    education = c("high school", "bachelor's", "master's"),
    stringsAsFactors = FALSE
  )
)

# Name the agents by their IDs
agents_named <- setNames(agents, sapply(agents, function(a) a$id))

# Find moderator ID (create_diverse_agents adds a moderator with ID "MOD")
moderator_id <- "MOD"

# Create a conversation flow object
flow_obj <- DesireBasedFlow$new(agents_named, moderator_id)

# Create the FocusGroup object
fg <- FocusGroup$new(
  topic = "Impact of Social Media on Mental Health",
  purpose = "To explore perspectives and experiences related to social media.",
  agents = agents_named,
  moderator_id = moderator_id,
  turn_taking_flow = flow_obj
)
```

## Use the transcript in other computations

The transcript data frame can be passed to ordinary R functions.

```{r custom_analysis, eval = FALSE}
# Extract conversation data
conv_data <- result$conversation

# Custom analysis example
word_counts <- sapply(conv_data$text, function(x) length(strsplit(x, "\\s+")[[1]]))
summary(word_counts)
```

# Interpretation

A generated transcript records model responses under a specified script,
persona set, and model configuration. It does not measure a population.
