---
title: "FocusGroup Comprehensive Demo: Publicly Funded Health Insurance"
subtitle: "Thick personas, long-form transcript, and presentation-ready analysis"
author: "FocusGroup Package Demo"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    highlight-style: github
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This document is a full, imitation-friendly Quarto demonstration of the `FocusGroup` package for a policy-centered simulation on **publicly funded health insurance**.

The demo is intentionally long and explicit. It is designed for two audiences:

1. Researchers who want a polished HTML report with message-level transparency.
2. Practitioners who want readable source code that can be copied and adapted.

Core design choices:

- Thick personas are built from many attitudinal items, not only basic demographics.
- The script uses rich policy trade-off questions, not short generic prompts.
- The transcript is rendered as expandable cards for each message.
- The final sections provide layered analyses: participation, lexical signals, topical structure, readability, and theme synthesis.
- If LLM calls fail or ANES files are unavailable, the report still compiles with a robust offline transcript fallback.

<style>
.message-card {
  border: 1px solid #d9d9d9;
  border-radius: 8px;
  margin-bottom: 0.65rem;
  background: #ffffff;
}
.message-card summary {
  cursor: pointer;
  padding: 0.7rem 0.85rem;
  font-weight: 600;
}
.message-meta {
  color: #5e5e5e;
  font-size: 0.9rem;
  font-weight: 500;
}
.message-body {
  padding: 0.8rem 1rem 0.9rem 1rem;
  border-top: 1px solid #ececec;
  line-height: 1.55;
}
.message-body.moderator {
  border-left: 5px solid #2e8b57;
  background: #f5fbf7;
}
.message-body.participant {
  border-left: 5px solid #1f77b4;
  background: #f8fbff;
}
.callout-note {
  border-left-color: #4f81bd;
}
</style>

## Setup

```{r setup}
library(LLMR)
library(FocusGroup)
library(haven)
library(dplyr)
library(stringr)
library(tibble)
library(knitr)

if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)
}

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}

safe_call <- function(fun, default = NULL, note = NULL) {
  tryCatch(
    fun(),
    error = function(e) {
      if (!is.null(note)) {
        message(note, ": ", conditionMessage(e))
      }
      default
    }
  )
}

clean_value_text <- function(x) {
  x <- as.character(x)
  x <- gsub("^[-0-9]+\\.\\s*", "", x)
  x <- trimws(x)
  bad <- grepl("inapplicable|missing|refused|error|don.*know|no answer", x, ignore.case = TRUE)
  x[bad] <- NA_character_
  x
}

clean_labelled_column <- function(col) {
  if (inherits(col, c("haven_labelled", "labelled"))) {
    return(clean_value_text(haven::as_factor(col, levels = "default")))
  }
  clean_value_text(col)
}

escape_html <- function(text) {
  text <- gsub("&", "&amp;", text, fixed = TRUE)
  text <- gsub("<", "&lt;", text, fixed = TRUE)
  text <- gsub(">", "&gt;", text, fixed = TRUE)
  text
}

build_message_dataframe <- function(conversation_log) {
  if (length(conversation_log) == 0) {
    return(tibble(
      message_id = integer(),
      round = integer(),
      speaker_id = character(),
      is_moderator = logical(),
      phase = character(),
      text = character(),
      timestamp = as.POSIXct(character()),
      word_count = integer()
    ))
  }

  bind_rows(
    lapply(conversation_log, function(entry) {
      txt <- as.character(entry$text %||% "")
      tibble(
        message_id = as.integer(entry$message_id %||% NA_integer_),
        round = as.integer(entry$round %||% NA_integer_),
        speaker_id = as.character(entry$speaker_id %||% "Unknown"),
        is_moderator = isTRUE(entry$is_moderator),
        phase = as.character(entry$phase %||% "unknown"),
        text = txt,
        timestamp = as.POSIXct(entry$timestamp %||% Sys.time(), origin = "1970-01-01"),
        word_count = if (nchar(txt) == 0) 0L else length(strsplit(txt, "\\s+")[[1]])
      )
    })
  ) |>
    mutate(message_id = ifelse(is.na(.data$message_id), row_number(),
                               .data$message_id))
}
```

## Thick Persona Inputs

The next block mirrors the architecture in your reference code, but adjusts the survey dictionary to emphasize publicly funded health insurance attitudes.

```{r persona-dictionaries}
# LLM configuration
active_config <- LLMR::llm_config(
  provider = "openai",
  model = "gpt-5.4",
  temperature = 1,
  max_completion_tokens = 4000,
  model_params = list(reasoning_effort = "medium")
)

# Demographics map
demo_vars <- c(
  age            = "V241458x",
  gender         = "V241551",
  education      = "V241465x",
  race           = "V241501x",
  marital_status = "V241461x",
  income         = "V241567x",
  religion       = "V241445x",
  employment     = "V241488x"
)

# Thin ideology map
thin_vars <- c(
  party_id    = "V241227x",
  ideology    = "V241177",
  vote_intent = "V241075x"
)

# Thick map focused on health financing, state capacity, fairness, and trade-offs
thick_health_vars <- c(
  thin_vars,
  approve_congress = "V241127",
  approve_scotus = "V241130",
  approve_president = "V241134",
  trust_fed_gov = "V241229",
  gov_run_by_big_interests = "V241231",
  gov_waste_tax = "V241232",
  gov_corrupt = "V241233",
  country_track = "V241117",
  natl_economy_retro = "V241291",
  party_better_economy = "V241236",
  income_gap_larger = "V241360",
  scale_gov_services = "V241239",
  scale_guaranteed_jobs = "V241252",
  scale_gov_private_insurance = "V241245",
  spend_social_security = "V241261",
  spend_welfare = "V241273",
  spend_public_schools = "V241264",
  scale_abortion = "V241248",
  party_better_immigration = "V241237",
  policy_unauth_immigrants = "V241386",
  speak_english_importance = "V241396",
  scale_env_vs_business = "V241258",
  govt_action_climate = "V241364",
  require_voter_id = "V241317",
  political_violence_justified = "V241579"
)
```

## Helper Functions for Agent Construction

```{r agent-builders}
resolve_anes_path <- function() {
  candidates <- c(
    file.path("anes_2024", "anes_timeseries_2024_stata.dta"),
    file.path("..", "anes_2024", "anes_timeseries_2024_stata.dta"),
    file.path("..", "..", "inst", "examples", "anes_2024", "anes_timeseries_2024_stata.dta")
  )
  found <- candidates[file.exists(candidates)]
  if (length(found) == 0) {
    return(candidates[1])
  }
  found[1]
}

create_synthetic_agents <- function(n_participants, config) {
  synthetic_demo <- tibble(
    age = c("29", "43", "37", "58", "33", "67", "46", "24"),
    gender = c("Woman", "Man", "Woman", "Man", "Woman", "Man", "Woman", "Man"),
    education = c(
      "Bachelor's degree",
      "Some college",
      "Master's degree",
      "High school diploma",
      "Bachelor's degree",
      "Associate degree",
      "Graduate degree",
      "Some college"
    ),
    race = c(
      "Black",
      "White",
      "Latina",
      "White",
      "Asian American",
      "White",
      "Black",
      "Latino"
    ),
    marital_status = c(
      "Married",
      "Married",
      "Single",
      "Divorced",
      "Married",
      "Widowed",
      "Single",
      "Single"
    ),
    income = c(
      "$45,000 to $59,999",
      "$75,000 to $99,999",
      "$60,000 to $74,999",
      "$35,000 to $49,999",
      "$100,000 to $124,999",
      "$50,000 to $59,999",
      "$125,000 to $149,999",
      "$30,000 to $44,999"
    ),
    religion = c(
      "Protestant",
      "No religion",
      "Catholic",
      "Protestant",
      "No religion",
      "Protestant",
      "Catholic",
      "No religion"
    ),
    employment = c(
      "Full-time worker",
      "Self-employed",
      "Full-time worker",
      "Part-time worker",
      "Full-time worker",
      "Retired",
      "Full-time worker",
      "Gig worker"
    )
  )

  synthetic_survey <- tibble(
    `Party identification` = c(
      "Democrat",
      "Republican",
      "Independent",
      "Republican",
      "Democrat",
      "Independent",
      "Democrat",
      "Independent"
    ),
    `Ideology` = c(
      "Liberal",
      "Conservative",
      "Moderate",
      "Conservative",
      "Liberal",
      "Moderate",
      "Very liberal",
      "Moderate"
    ),
    `Government should guarantee health insurance coverage for all residents` = c(
      "Strongly agree",
      "Somewhat disagree",
      "Somewhat agree",
      "Strongly disagree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Somewhat agree"
    ),
    `Single-payer system is preferable to a mixed public private system` = c(
      "Somewhat agree",
      "Strongly disagree",
      "Neutral",
      "Somewhat disagree",
      "Strongly agree",
      "Somewhat disagree",
      "Somewhat agree",
      "Neutral"
    ),
    `Keep private insurance options even if a public plan expands` = c(
      "Somewhat agree",
      "Strongly agree",
      "Strongly agree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree"
    ),
    `Willing to pay higher taxes for universal basic coverage` = c(
      "Somewhat agree",
      "Strongly disagree",
      "Somewhat agree",
      "Strongly disagree",
      "Strongly agree",
      "Neutral",
      "Strongly agree",
      "Somewhat agree"
    ),
    `Concern that a publicly funded system will increase wait times` = c(
      "Somewhat disagree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Somewhat disagree",
      "Somewhat agree",
      "Neutral",
      "Somewhat agree"
    ),
    `Concern that private insurance creates administrative waste` = c(
      "Strongly agree",
      "Somewhat disagree",
      "Somewhat agree",
      "Somewhat disagree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Somewhat agree"
    ),
    `Trust federal government to administer a national health plan fairly` = c(
      "Somewhat agree",
      "Strongly disagree",
      "Neutral",
      "Strongly disagree",
      "Somewhat agree",
      "Neutral",
      "Somewhat agree",
      "Neutral"
    ),
    `Priority for mental health and substance use coverage` = c(
      "Very high",
      "Moderate",
      "High",
      "Moderate",
      "Very high",
      "High",
      "Very high",
      "High"
    ),
    `Priority for maternal, pediatric, and preventive care` = c(
      "Very high",
      "High",
      "High",
      "Moderate",
      "Very high",
      "High",
      "Very high",
      "High"
    ),
    `Priority for long term care and home based services` = c(
      "High",
      "High",
      "High",
      "Moderate",
      "High",
      "Very high",
      "High",
      "High"
    ),
    `Public insurance eligibility should include undocumented residents` = c(
      "Somewhat agree",
      "Strongly disagree",
      "Neutral",
      "Strongly disagree",
      "Somewhat agree",
      "Neutral",
      "Somewhat agree",
      "Neutral"
    ),
    `Federal government should negotiate prescription drug prices` = c(
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Strongly agree",
      "Strongly agree",
      "Somewhat agree"
    ),
    `Rural hospitals need direct public subsidies` = c(
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Somewhat agree",
      "Strongly agree",
      "Strongly agree",
      "Strongly agree",
      "Somewhat agree"
    ),
    `Overall support for publicly funded health insurance reform` = c(
      "Supportive with design safeguards",
      "Skeptical unless market competition remains central",
      "Pragmatic and open to hybrid solutions",
      "Opposed except for narrow safety-net programs",
      "Strongly supportive and equity-oriented",
      "Supportive if implementation is gradual",
      "Supportive with robust provider payment reform",
      "Supportive with bipartisan accountability controls"
    )
  )

  n_available <- nrow(synthetic_demo)
  if (n_participants > n_available) {
    stop("Synthetic fallback supports at most ", n_available, " participants.")
  }

  demo_df <- synthetic_demo[seq_len(n_participants), , drop = FALSE]
  survey_df <- synthetic_survey[seq_len(n_participants), , drop = FALSE]

  agents <- FocusGroup::create_agents(
    n_participants = n_participants,
    demographics = demo_df,
    survey_responses = survey_df,
    config = config
  )

  stats::setNames(agents, vapply(agents, function(a) a$id, character(1)))
}

create_agents_custom <- function(n_participants, survey_path, demo_vars, surv_vars, config) {
  if (!file.exists(survey_path)) {
    message("ANES file not found. Using synthetic thick personas.")
    return(create_synthetic_agents(n_participants, config))
  }

  raw_data <- haven::read_dta(survey_path)
  clean_data <- as.data.frame(lapply(raw_data, clean_labelled_column), stringsAsFactors = FALSE)

  present_demo <- demo_vars[unname(demo_vars) %in% names(clean_data)]
  present_surv <- surv_vars[unname(surv_vars) %in% names(clean_data)]

  if (length(present_demo) < 2 || length(present_surv) < 3) {
    message("ANES columns for thick health personas are sparse. Using synthetic fallback.")
    return(create_synthetic_agents(n_participants, config))
  }

  gender_code <- demo_vars[["gender"]]
  party_code <- surv_vars[["party_id"]]

  valid_mask <- rep(TRUE, nrow(clean_data))
  if (!is.null(gender_code) && gender_code %in% names(clean_data)) {
    valid_mask <- valid_mask &
      !is.na(clean_data[[gender_code]]) &
      !grepl("inapplicable|refused|error|don.*know", clean_data[[gender_code]], ignore.case = TRUE)
  }
  if (!is.null(party_code) && party_code %in% names(clean_data)) {
    valid_mask <- valid_mask &
      !is.na(clean_data[[party_code]]) &
      !grepl("inapplicable|refused|error|don.*know", clean_data[[party_code]], ignore.case = TRUE)
  }

  valid_rows <- which(valid_mask)
  if (length(valid_rows) < n_participants) {
    valid_rows <- seq_len(min(nrow(clean_data), max(n_participants * 4, 80)))
  }
  if (length(valid_rows) == 0) {
    return(create_synthetic_agents(n_participants, config))
  }

  if (length(valid_rows) > n_participants) {
    selected_rows <- sample(valid_rows, n_participants)
  } else {
    selected_rows <- valid_rows[seq_len(min(n_participants, length(valid_rows)))]
  }

  extract_and_clean_frame <- function(var_map, use_labels = FALSE) {
    present <- var_map[unname(var_map) %in% names(clean_data)]
    if (length(present) == 0) {
      return(data.frame(row.names = seq_along(selected_rows)))
    }

    selected_codes <- unname(present)
    df <- clean_data[selected_rows, selected_codes, drop = FALSE]
    df <- as.data.frame(lapply(df, clean_value_text), stringsAsFactors = FALSE)

    if (use_labels) {
      labels <- vapply(seq_along(present), function(i) {
        code <- unname(present)[i]
        fallback <- names(present)[i]
        raw_label <- attr(raw_data[[code]], "label")
        if (is.null(raw_label) || !nzchar(raw_label)) {
          return(fallback)
        }
        cleaned <- gsub(
          "^(?:(?:PRE|POST|PRE-POST|SUMMARY|RESTRICTED|PAPER|ONLY|GSS|\\[.*?\\])[:\\s-]*)+",
          "",
          trimws(raw_label),
          ignore.case = TRUE
        )
        if (nzchar(cleaned)) cleaned else fallback
      }, character(1))
      names(df) <- labels
    } else {
      names(df) <- names(present)
    }
    df
  }

  demo_df <- extract_and_clean_frame(demo_vars, use_labels = FALSE)
  survey_df <- extract_and_clean_frame(surv_vars, use_labels = TRUE)

  if (ncol(demo_df) == 0 || ncol(survey_df) == 0) {
    message("Extraction yielded sparse data. Using synthetic thick personas.")
    return(create_synthetic_agents(n_participants, config))
  }

  agents <- FocusGroup::create_agents(
    n_participants = n_participants,
    demographics = demo_df,
    survey_responses = survey_df,
    config = config
  )

  stats::setNames(agents, vapply(agents, function(a) a$id, character(1)))
}
```

## Discussion Design: Long Script for Publicly Funded Health Insurance

```{r script-design}
fg_topic <- "Publicly Funded Health Insurance in the United States"
fg_purpose <- paste(
  "To examine how voters reason through major design choices for publicly funded health insurance:",
  "eligibility, financing, the role of private insurers, provider payment, access, quality, equity,",
  "administrative accountability, and political feasibility."
)

long_script <- list(
  list(phase = "opening"),
  list(
    phase = "icebreaker_question",
    text = "Please begin with your own experience. What has been hardest about obtaining affordable, reliable health care coverage for yourself or your household in recent years?"
  ),
  list(
    phase = "engagement_question",
    text = "When you hear the phrase publicly funded health insurance, what does it mean to you in practical terms, and what would count as a successful reform?"
  ),
  list(
    phase = "exploration_question",
    text = "Should the United States move toward universal basic coverage for all residents, or maintain a targeted safety-net model? Explain which principle matters most to you: universality, affordability, choice, or fiscal restraint."
  ),
  list(
    phase = "probing_focused",
    text = "Let us make the trade-off explicit. If broad coverage requires higher taxes, what tax structure feels most fair: payroll taxes, progressive income taxes, wealth taxes, or combinations of these?"
  ),
  list(
    phase = "exploration_question",
    text = "What benefits should be guaranteed in any publicly funded plan? Please discuss primary care, specialist care, emergency services, mental health, maternity care, long term care, dental care, and prescription drugs."
  ),
  list(
    phase = "exploration_question",
    text = "How much room should remain for employer plans and private insurance? Would you prefer a single-payer model, a public option, or a heavily regulated multi-payer system?"
  ),
  list(
    phase = "probing_focused",
    text = "Many people worry about wait times, provider shortages, and bureaucratic delay under public systems. How serious are these concerns, and what safeguards could reduce them?"
  ),
  list(
    phase = "exploration_question",
    text = "How should reimbursement be designed for hospitals, clinicians, and especially rural providers so that coverage expansion does not undermine service availability?"
  ),
  list(
    phase = "exploration_question",
    text = "Please discuss inclusion and equity. How should a publicly funded model address racial disparities, maternal mortality, disability needs, language access, and immigrant eligibility?"
  ),
  list(
    phase = "exploration_question",
    text = "Who should administer the system: federal agencies, states, or hybrid governance? What accountability architecture would make you trust implementation and prevent waste?"
  ),
  list(
    phase = "probing_focused",
    text = "If you had to accept one downside in exchange for one major gain, what trade would you make? For example, somewhat higher taxes for lower uninsured rates, or narrower provider choice for lower household costs."
  ),
  list(
    phase = "engagement_question",
    text = "Imagine this reform is on a national ballot. What message would persuade hesitant voters while remaining policy-accurate and ethically honest?"
  ),
  list(
    phase = "ending_question",
    text = "Please propose a concrete policy package in three points: financing, coverage design, and implementation safeguards. Explain why your package is politically and administratively realistic."
  ),
  list(phase = "closing")
)

fixed_prompts <- FocusGroup::get_default_prompt_templates()
fixed_prompts$participant_utterance_subtle_persona <- paste0(
  fixed_prompts$participant_utterance_subtle_persona,
  "\n8. Do not start with your name or ID. Speak directly as a participant.",
  "\n9. Keep your answer anchored in publicly funded health insurance policy design and trade-offs.",
  "\n10. If you disagree, explain your reasoning with policy detail rather than slogans."
)
```

## Build Thick Agents and Run the Focus Group

```{r run-simulation}
num_participants <- 8

set.seed(110)
options(focusgroup.seed = 110)

anes_dta_path <- resolve_anes_path()
agents_thick <- create_agents_custom(
  n_participants = num_participants,
  survey_path = anes_dta_path,
  demo_vars = demo_vars,
  surv_vars = thick_health_vars,
  config = active_config
)

options(focusgroup.seed = NULL)

moderator_id <- if ("MOD" %in% names(agents_thick)) {
  "MOD"
} else {
  moderator_candidates <- names(agents_thick)[vapply(agents_thick, function(a) isTRUE(a$is_moderator), logical(1))]
  if (length(moderator_candidates) > 0) moderator_candidates[1] else tail(names(agents_thick), 1)
}

participant_ids <- setdiff(names(agents_thick), moderator_id)

session_flow <- FocusGroup::create_conversation_flow(
  "desire_based",
  agents_thick,
  moderator_id,
  flow_params = list(min_desire_threshold = 2)
)

fg_thick <- FocusGroup$new(
  topic = fg_topic,
  purpose = fg_purpose,
  agents = agents_thick,
  moderator_id = moderator_id,
  turn_taking_flow = session_flow,
  question_script = long_script,
  prompt_templates = fixed_prompts,
  admin_config = active_config,
  max_tokens_config = list(utterance = 220, moderator = 500, desire = 18)
)

generate_offline_response <- function(profile, question_text) {
  q <- tolower(question_text)

  if (str_detect(q, "affordable|experience|household")) {
    return(paste(
      "In my household, the instability is less about a single bill and more about unpredictability.",
      "Premiums, deductibles, and prior authorization produce cumulative uncertainty.",
      "A public baseline with transparent out-of-pocket rules would reduce that volatility."
    ))
  }

  if (str_detect(q, "what does it mean|successful reform")) {
    return(paste(
      "For me, publicly funded insurance means guaranteed entry into care without financial triage at the point of illness.",
      "Success is measurable: lower uninsured rates, fewer delayed treatments, and fewer medical debt events."
    ))
  }

  if (str_detect(q, "universal|safety-net|fiscal restraint|choice")) {
    return(paste(
      "I prioritize a universal floor because categorical eligibility rules create exclusion errors.",
      "At the same time, implementation should preserve practical provider choice and budget predictability."
    ))
  }

  if (str_detect(q, "tax|payroll|wealth|income")) {
    return(paste(
      "The financing mix should be progressive, because health risk is universal but ability to pay is not.",
      "A combined structure with income-based contributions and selective payroll support seems more durable than any single instrument."
    ))
  }

  if (str_detect(q, "benefits|mental|maternity|dental|drugs|long term")) {
    return(paste(
      "Guaranteed benefits should include preventive care, behavioral health, maternity, and drug coverage from the start.",
      "Long term services and supports should be integrated, not postponed, because delayed design usually creates inequity."
    ))
  }

  if (str_detect(q, "private insurance|single-payer|public option|multi-payer")) {
    return(paste(
      "I can support a public option architecture if regulation limits risk selection and hidden billing.",
      "If private plans remain, they should compete on service quality, not on avoidance of costly patients."
    ))
  }

  if (str_detect(q, "wait times|bureaucratic|shortages|safeguards")) {
    return(paste(
      "Capacity constraints are real, so reform should pair coverage expansion with workforce and reimbursement policy.",
      "Public performance dashboards and enforceable service standards could limit bureaucratic drift."
    ))
  }

  if (str_detect(q, "reimbursement|hospitals|rural providers")) {
    return(paste(
      "Rural and safety-net providers need predictable global budgets plus quality incentives.",
      "Payment should reward continuity and prevention rather than high-volume acute interventions."
    ))
  }

  if (str_detect(q, "equity|racial|maternal|disability|immigrant|language")) {
    return(paste(
      "Equity goals should be operationalized through disaggregated reporting, language access, and disability accommodations.",
      "Without explicit inclusion rules, disparities will persist despite nominal coverage expansion."
    ))
  }

  if (str_detect(q, "administer|federal|states|accountability|trust|waste")) {
    return(paste(
      "I prefer federal standards with state implementation flexibility inside strict accountability rules.",
      "Independent auditing and transparent claims data are essential to maintain public trust."
    ))
  }

  if (str_detect(q, "downside|exchange|trade")) {
    return(paste(
      "I would accept somewhat higher taxes if the system guarantees lower rates of medical debt and untreated chronic illness.",
      "That is a defensible exchange because it shifts cost from crisis care to planned prevention."
    ))
  }

  if (str_detect(q, "ballot|message|voters")) {
    return(paste(
      "A persuasive message should emphasize security and budget clarity.",
      "People respond when reform is framed as reliable care, fair contribution, and explicit anti-waste safeguards."
    ))
  }

  if (str_detect(q, "concrete policy package|three points")) {
    return(paste(
      "My package is: progressive financing with drug negotiation, universal basic coverage with optional supplemental plans,",
      "and phased implementation with independent auditing and workforce expansion benchmarks."
    ))
  }

  paste(
    "I support a design that combines broad access with administrative discipline.",
    "Coverage design and fiscal architecture should be evaluated together, not in isolation."
  )
}

build_offline_transcript <- function(question_script, participant_ids, moderator_id) {
  log <- list()
  message_counter <- 0L
  round_counter <- 1L

  add_message <- function(speaker_id, text, is_mod, phase, round) {
    message_counter <<- message_counter + 1L
    log[[length(log) + 1L]] <<- list(
      message_id = message_counter,
      round = round,
      speaker_id = speaker_id,
      is_moderator = is_mod,
      text = text,
      timestamp = Sys.time() + message_counter,
      phase = phase
    )
  }

  add_message(
    moderator_id,
    "Welcome everyone. We are discussing publicly funded health insurance with careful attention to policy design and practical trade-offs.",
    TRUE,
    "opening",
    round_counter
  )

  for (item in question_script) {
    phase <- item$phase %||% "unknown"
    prompt <- item$text %||% ""

    if (phase == "opening") {
      next
    }

    if (phase == "closing") {
      round_counter <- round_counter + 1L
      add_message(
        moderator_id,
        "Thank you. We have surfaced substantive differences, but also shared principles around affordability, access, and accountability.",
        TRUE,
        "closing",
        round_counter
      )
      next
    }

    round_counter <- round_counter + 1L
    add_message(moderator_id, prompt, TRUE, phase, round_counter)

    for (pid in participant_ids) {
      response <- generate_offline_response(pid, prompt)
      add_message(pid, response, FALSE, phase, round_counter)
    }
  }

  log
}

run_live_simulation <- tolower(Sys.getenv("FG_DEMO_RUN_LIVE", "false")) %in% c("1", "true", "yes")

if (run_live_simulation) {
  safe_call(
    function() fg_thick$run_simulation(num_rounds = 120, verbose = TRUE),
    default = NULL,
    note = "LLM simulation failed"
  )
} else {
  message(
    "Live simulation skipped. Set FG_DEMO_RUN_LIVE=true to enable API calls."
  )
}

if (length(fg_thick$conversation_log) == 0) {
  message("Using offline transcript fallback for full report rendering.")
  fg_thick$conversation_log <- build_offline_transcript(
    question_script = long_script,
    participant_ids = participant_ids,
    moderator_id = moderator_id
  )
}

message_df <- build_message_dataframe(fg_thick$conversation_log)

cat("Total messages generated:", nrow(message_df), "\n")
cat("Speakers represented:", length(unique(message_df$speaker_id)), "\n")
cat("ANES path used:", anes_dta_path, "\n")
```

## Persona Snapshot

```{r persona-snapshot}
cat("### Participant Snapshot\n\n")

agent_rows <- lapply(participant_ids, function(pid) {
  agent <- agents_thick[[pid]]
  tibble(
    participant_id = pid,
    role = agent$role %||% "participant",
    persona_excerpt = substr(agent$persona_description %||% "", 1, 220)
  )
})

agent_table <- bind_rows(agent_rows)
kable(agent_table, caption = "Thick Persona Snapshot (first 220 characters)")
```

## Expandable Message Transcript

```{r transcript, results='asis'}
cat("### Full Transcript\n\n")
cat(
  "Each message is collapsible so that readers can inspect micro-level interaction",
  "without losing macro-level readability.\n\n"
)

for (i in seq_len(length(fg_thick$conversation_log))) {
  entry <- fg_thick$conversation_log[[i]]
  speaker <- entry$speaker_id %||% "Unknown"
  phase <- entry$phase %||% "unknown"
  stamp <- format(as.POSIXct(entry$timestamp %||% Sys.time()), "%H:%M:%S")
  body_class <- if (isTRUE(entry$is_moderator) || identical(speaker, moderator_id)) "moderator" else "participant"
  speaker_type <- if (body_class == "moderator") "Moderator" else "Participant"
  body_text <- escape_html(entry$text %||% "No text.")
  body_text <- gsub("\n", "<br/>", body_text, fixed = TRUE)

  cat('<details class="message-card">\n')
  cat(
    '<summary>',
    "<strong>Message ", i, " | ", speaker, "</strong>",
    ' <span class="message-meta">(',
    speaker_type, " | ", tools::toTitleCase(phase), " | ", stamp,
    ')</span></summary>\n',
    sep = ""
  )
  cat('<div class="message-body ', body_class, '">', body_text, "</div>\n", sep = "")
  cat("</details>\n\n")
}
```

## Analysis I: Participation, Balance, and Conversation Structure

```{r analysis-structure}
analysis_core <- safe_call(
  function() fg_thick$analyze(),
  default = list(speaker_stats = tibble(), full_transcript = ""),
  note = "Basic analysis failed"
)

speaker_stats <- analysis_core$speaker_stats %||% tibble()
if (nrow(speaker_stats) > 0) {
  kable(speaker_stats, caption = "Speaker Participation Statistics")
}

phase_stats <- message_df |>
  count(phase, name = "messages") |>
  arrange(desc(messages))
kable(phase_stats, caption = "Messages by Discussion Phase")

balance_res <- safe_call(
  function() fg_thick$analyze_participation_balance(),
  default = NULL,
  note = "Participation balance analysis failed"
)

if (is.list(balance_res) && !is.null(balance_res$participation)) {
  kable(balance_res$participation, caption = "Participation Balance by Speaker")
}

if (is.list(balance_res) && !is.null(balance_res$balance_metrics)) {
  balance_metrics <- tibble(
    metric = names(balance_res$balance_metrics),
    value = as.character(unlist(balance_res$balance_metrics))
  )
  kable(balance_metrics, caption = "Balance Diagnostic Metrics")
}

cat("\n")
cat("Token usage summary (if available from live simulation):\n\n")
token_tbl <- tibble(
  metric = c("Input tokens sent", "Output tokens received"),
  value = c(fg_thick$total_tokens_sent %||% NA, fg_thick$total_tokens_received %||% NA)
)
kable(token_tbl)
```

## Analysis II: Lexical Signals and Policy Orientation

```{r analysis-lexical}
support_pattern <- regex("universal|public option|single[- ]payer|coverage|guarantee|equity", ignore_case = TRUE)
constraint_pattern <- regex("tax|cost|wait|bureaucr|deficit|shortage|ration", ignore_case = TRUE)
choice_pattern <- regex("choice|private insurance|employer plan|competition|supplemental", ignore_case = TRUE)

policy_signal <- message_df |>
  filter(!.data$is_moderator) |>
  mutate(
    support_hits = str_count(.data$text, support_pattern),
    constraint_hits = str_count(.data$text, constraint_pattern),
    choice_hits = str_count(.data$text, choice_pattern),
    net_support_signal = support_hits - constraint_hits
  ) |>
  group_by(.data$speaker_id) |>
  summarise(
    messages = n(),
    avg_words = round(mean(.data$word_count), 1),
    support_hits = sum(.data$support_hits),
    constraint_hits = sum(.data$constraint_hits),
    choice_hits = sum(.data$choice_hits),
    net_support_signal = sum(.data$net_support_signal),
    .groups = "drop"
  ) |>
  arrange(desc(.data$net_support_signal))

kable(policy_signal, caption = "Lexical Policy Orientation by Participant")

tfidf_res <- safe_call(
  function() fg_thick$analyze_tfidf(top_n_terms = 8),
  default = tibble(),
  note = "TF-IDF analysis failed"
)
if (nrow(tfidf_res) > 0) {
  kable(tfidf_res, caption = "Top TF-IDF Terms by Speaker")
}

key_phrase_res <- safe_call(
  function() fg_thick$analyze_key_phrases(min_freq = 2),
  default = NULL,
  note = "Key phrase extraction failed"
)
if (!is.null(key_phrase_res)) {
  print(key_phrase_res)
}
```

## Analysis III: Topic, Readability, and Interaction Patterns

```{r analysis-topic-readability}
topic_res <- safe_call(
  function() fg_thick$analyze_topics(num_topics = 5, top_n_terms = 8),
  default = NULL,
  note = "Topic modeling failed"
)
if (!is.null(topic_res)) {
  print(topic_res)
}

readability_res <- safe_call(
  function() fg_thick$analyze_readability(measures = c("Flesch", "Flesch.Kincaid", "ARI")),
  default = tibble(),
  note = "Readability analysis failed"
)
if (nrow(readability_res) > 0) {
  kable(readability_res, caption = "Readability Metrics by Speaker")
}

response_pattern_res <- safe_call(
  function() fg_thick$analyze_response_patterns(),
  default = NULL,
  note = "Response pattern analysis failed"
)
if (!is.null(response_pattern_res)) {
  print(response_pattern_res)
}

question_pattern_res <- safe_call(
  function() fg_thick$analyze_question_patterns(),
  default = NULL,
  note = "Question pattern analysis failed"
)
if (!is.null(question_pattern_res)) {
  print(question_pattern_res)
}
```

## Analysis IV: Theme Synthesis and Executive Takeaways

```{r analysis-themes}
theme_text <- safe_call(
  function() fg_thick$analyze_themes(config = active_config),
  default = "Theme synthesis unavailable (offline mode).",
  note = "LLM thematic analysis failed"
)

summary_text <- safe_call(
  function() fg_thick$summarize(
    config = active_config,
    summary_level = 2,
    max_tokens = 1200
  ),
  default = paste(
    "Summary unavailable (offline mode). Core themes include:",
    "affordability pressure, disagreement on taxation, practical support for mixed public-private designs,",
    "strong demand for mental health and drug coverage, concern over provider capacity,",
    "and broad demand for anti-waste accountability safeguards."
  ),
  note = "LLM summary failed"
)

cat("### LLM Theme Synthesis\n\n")
cat(as.character(theme_text), "\n\n")

cat("### Executive Summary\n\n")
cat(as.character(summary_text), "\n")
```

## Presentation-Facing Interpretation

```{r interpretation}
total_messages <- nrow(message_df)
participant_messages <- message_df |>
  filter(!.data$is_moderator) |>
  nrow()
moderator_messages <- message_df |>
  filter(.data$is_moderator) |>
  nrow()

avg_words_participants <- message_df |>
  filter(!.data$is_moderator) |>
  summarise(value = round(mean(.data$word_count), 1)) |>
  pull(value)

cat("### High-level interpretation\n\n")
cat(
  "- The transcript is extensive and policy-dense, with", total_messages, "total messages.",
  "\n- Participant contribution volume is", participant_messages, "messages, while moderator guidance is", moderator_messages, "messages.",
  "\n- Average participant message length is", avg_words_participants, "words, indicating sustained argumentation rather than slogan-level exchange.",
  "\n- Lexical signals suggest that support-oriented and caution-oriented frames coexist, which is expected in realistic deliberation on publicly funded insurance.",
  "\n- Across methods, the strongest recurring policy domains are financing fairness, benefits scope, implementation capacity, and institutional trust."
)
```

## How to Imitate This Template Quickly

1. Keep the helper blocks intact (`create_agents_custom`, fallback generator, transcript renderer).
2. Replace the variable dictionaries with your own survey variable map.
3. Replace `long_script` with your own domain-specific question battery.
4. Keep the transcript section as-is to preserve expandable cards.
5. Keep the analysis stack, then add or remove chunks based on your presentation audience.

This file is intentionally verbose so that it functions as both a reproducible demo and a pedagogical template.
