---
title: "FocusGroup Comprehensive Demo with DeepSeek"
subtitle: "Publicly funded health insurance with mixed partisan thick personas"
author: "FocusGroup Package Demo"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    highlight-style: github
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This file is intentionally live only. It does not silently fabricate a transcript.  
It runs API calls through DeepSeek and then performs a final direct LLMR analysis on:

1. Full participant personas
2. Full transcript
3. Metadata for party and education

The topic is publicly funded health insurance. The script is long, and participant responses are forced to be longer than one sentence.

<style>
.turn-card {
  border: 1px solid #d9d9d9;
  border-radius: 8px;
  margin-bottom: 0.65rem;
  background: #ffffff;
}
.turn-card summary {
  cursor: pointer;
  padding: 0.7rem 0.85rem;
  font-weight: 600;
}
.turn-meta {
  color: #5e5e5e;
  font-size: 0.9rem;
  font-weight: 500;
}
.turn-body {
  padding: 0.8rem 1rem 0.9rem 1rem;
  border-top: 1px solid #ececec;
  line-height: 1.6;
}
.turn-body.moderator {
  border-left: 5px solid #2e8b57;
  background: #f5fbf7;
}
.turn-body.participant {
  border-left: 5px solid #1f77b4;
  background: #f8fbff;
}
</style>

## Setup

```{r setup}
library(LLMR)
library(FocusGroup)
library(haven)
library(dplyr)
library(stringr)
library(tibble)
library(knitr)

`%||%` <- function(x, y) if (is.null(x) || length(x) == 0) y else x

clean_value_text <- function(x) {
  x <- as.character(x)
  x <- gsub("^[-0-9]+\\.\\s*", "", x)
  x <- trimws(x)
  bad <- grepl("inapplicable|missing|refused|error|don.*know|no answer", x, ignore.case = TRUE)
  x[bad] <- NA_character_
  x
}

clean_labelled_column <- function(col) {
  if (inherits(col, c("haven_labelled", "labelled"))) {
    return(clean_value_text(haven::as_factor(col, levels = "default")))
  }
  clean_value_text(col)
}

escape_html <- function(text) {
  text <- gsub("&", "&amp;", text, fixed = TRUE)
  text <- gsub("<", "&lt;", text, fixed = TRUE)
  text <- gsub(">", "&gt;", text, fixed = TRUE)
  text
}

unwrap_llm_text <- function(resp) {
  if (!is.null(resp$text)) return(as.character(resp$text))
  as.character(resp)
}

build_turn_dataframe <- function(conversation_log) {
  bind_rows(lapply(conversation_log, function(entry) {
    txt <- as.character(entry$text %||% "")
    tibble(
      turn = as.integer(entry$turn %||% NA_integer_),
      speaker_id = as.character(entry$speaker_id %||% "Unknown"),
      is_moderator = isTRUE(entry$is_moderator),
      phase = as.character(entry$phase %||% "unknown"),
      text = txt,
      timestamp = as.POSIXct(entry$timestamp %||% Sys.time(), origin = "1970-01-01"),
      word_count = if (nchar(txt) == 0) 0L else length(strsplit(txt, "\\s+")[[1]])
    )
  })) |>
    mutate(turn = ifelse(is.na(.data$turn), row_number(), .data$turn))
}
```

## DeepSeek Configuration

```{r llm-config}
cfg_agents <- LLMR::llm_config(
  provider = "deepseek",
  model = "deepseek-v4-pro",
  thinking = "enabled",
  reasoning_effort = "high",
  temperature = 0.9,
  max_tokens = 900
)

cfg_admin <- LLMR::llm_config(
  provider = "deepseek",
  model = "deepseek-v4-pro",
  thinking = "enabled",
  reasoning_effort = "high",
  temperature = 0.8,
  max_tokens = 2200
)
```

## Thick Dictionaries and Sampling Rules

```{r dictionaries}
demo_vars <- c(
  age            = "V241458x",
  gender         = "V241551",
  education      = "V241465x",
  race           = "V241501x",
  marital_status = "V241461x",
  income         = "V241567x",
  religion       = "V241445x",
  employment     = "V241488x"
)

thin_vars <- c(
  party_id    = "V241227x",
  ideology    = "V241177",
  vote_intent = "V241075x"
)

thick_health_vars <- c(
  thin_vars,
  approve_congress = "V241127",
  approve_scotus = "V241130",
  approve_president = "V241134",
  trust_fed_gov = "V241229",
  gov_run_by_big_interests = "V241231",
  gov_waste_tax = "V241232",
  gov_corrupt = "V241233",
  country_track = "V241117",
  natl_economy_retro = "V241291",
  party_better_economy = "V241236",
  income_gap_larger = "V241360",
  scale_gov_services = "V241239",
  scale_guaranteed_jobs = "V241252",
  scale_gov_private_insurance = "V241245",
  spend_social_security = "V241261",
  spend_welfare = "V241273",
  spend_public_schools = "V241264",
  scale_abortion = "V241248",
  party_better_immigration = "V241237",
  policy_unauth_immigrants = "V241386",
  speak_english_importance = "V241396",
  scale_env_vs_business = "V241258",
  govt_action_climate = "V241364",
  require_voter_id = "V241317",
  political_violence_justified = "V241579"
)

resolve_anes_path <- function() {
  candidates <- c(
    file.path("anes_2024", "anes_timeseries_2024_stata.dta"),
    file.path("..", "anes_2024", "anes_timeseries_2024_stata.dta"),
    file.path("..", "..", "inst", "examples", "anes_2024", "anes_timeseries_2024_stata.dta")
  )
  found <- candidates[file.exists(candidates)]
  if (length(found) == 0) stop("ANES .dta file not found in expected paths.")
  found[1]
}

party_bucket <- function(x) {
  x <- tolower(x %||% "")
  if (grepl("republican", x)) return("Republican")
  if (grepl("democrat", x)) return("Democrat")
  if (grepl("independent|non-partisan|no party", x)) return("Independent")
  "Other"
}

sample_mixed_rows <- function(clean_data, party_code, n_participants = 8) {
  party_raw <- clean_data[[party_code]]
  party_grp <- vapply(party_raw, party_bucket, character(1))
  valid_rows <- which(!is.na(party_raw))
  if (length(valid_rows) < n_participants) stop("Not enough valid party rows for requested participant count.")

  targets <- c(Democrat = 3, Republican = 3, Independent = 2)
  selected <- integer(0)
  for (grp in names(targets)) {
    idx <- intersect(valid_rows, which(party_grp == grp))
    take <- min(length(idx), targets[[grp]])
    if (take > 0) selected <- c(selected, sample(idx, take))
  }

  if (length(selected) < n_participants) {
    remainder <- setdiff(valid_rows, selected)
    need <- n_participants - length(selected)
    selected <- c(selected, sample(remainder, need))
  }
  selected
}

extract_frame <- function(raw_data, clean_data, rows, var_map, use_labels = FALSE) {
  present <- var_map[unname(var_map) %in% names(clean_data)]
  if (length(present) == 0) stop("No requested variables found in ANES file.")

  codes <- unname(present)
  df <- clean_data[rows, codes, drop = FALSE]
  df <- as.data.frame(lapply(df, clean_value_text), stringsAsFactors = FALSE)

  if (use_labels) {
    labs <- vapply(seq_along(present), function(i) {
      code <- unname(present)[i]
      fallback <- names(present)[i]
      lbl <- attr(raw_data[[code]], "label")
      if (is.null(lbl) || !nzchar(lbl)) return(fallback)
      lbl <- trimws(gsub(
        "^(?:(?:PRE|POST|PRE-POST|SUMMARY|RESTRICTED|PAPER|ONLY|GSS|\\[.*?\\])[:\\s-]*)+",
        "",
        lbl,
        ignore.case = TRUE
      ))
      if (nzchar(lbl)) lbl else fallback
    }, character(1))
    names(df) <- labs
  } else {
    names(df) <- names(present)
  }
  df
}
```

## Build Mixed Partisan Thick Personas

```{r create-agents}
set.seed(2026)
options(focusgroup.seed = 2026)

anes_path <- resolve_anes_path()
raw_data <- haven::read_dta(anes_path)
clean_data <- as.data.frame(lapply(raw_data, clean_labelled_column), stringsAsFactors = FALSE)

required_codes <- unique(c(unname(demo_vars), unname(thick_health_vars)))
missing_codes <- setdiff(required_codes, names(clean_data))
if (length(missing_codes) > 0) {
  stop("ANES file is missing required variables: ", paste(missing_codes[1:min(20, length(missing_codes))], collapse = ", "))
}

n_participants <- 8
selected_rows <- sample_mixed_rows(clean_data, party_code = thin_vars[["party_id"]], n_participants = n_participants)

demo_df <- extract_frame(raw_data, clean_data, selected_rows, demo_vars, use_labels = FALSE)
survey_df <- extract_frame(raw_data, clean_data, selected_rows, thick_health_vars, use_labels = TRUE)

agents <- FocusGroup::create_diverse_agents(
  n_participants = n_participants,
  demographics = demo_df,
  survey_responses = survey_df,
  llm_config = cfg_agents
)

for (i in seq_along(agents)) {
  if (!agents[[i]]$is_moderator && length(agents[[i]]$survey_responses) > 0) {
    agents[[i]]$persona_description <- paste(
      agents[[i]]$persona_description,
      FocusGroup::format_survey_responses(agents[[i]]$survey_responses),
      sep = "\n\n"
    )
  }
}

agents_named <- stats::setNames(agents, vapply(agents, function(a) a$id, character(1)))
moderator_id <- if ("MOD" %in% names(agents_named)) "MOD" else tail(names(agents_named), 1)
participant_ids <- setdiff(names(agents_named), moderator_id)

participant_meta <- tibble(
  row_id = selected_rows,
  participant_id = participant_ids[seq_len(n_participants)],
  party_group = vapply(clean_data[[thin_vars[["party_id"]]]][selected_rows], party_bucket, character(1)),
  education = clean_data[[demo_vars[["education"]]]][selected_rows] %||% NA_character_,
  ideology = clean_data[[thin_vars[["ideology"]]]][selected_rows] %||% NA_character_
)

options(focusgroup.seed = NULL)
set.seed(NULL)
```

## Long Script and Long-Answer Prompt Controls

```{r script-and-prompts}
fg_topic <- "Publicly funded health insurance in the United States"
fg_purpose <- paste(
  "To examine policy trade-offs in publicly funded health insurance, with explicit attention to",
  "financing, eligibility, administrative design, private-public boundaries, equity, and political feasibility."
)

script <- list(
  list(phase = "opening"),
  list(phase = "icebreaker_question", text = "Briefly describe one concrete challenge your household has faced in paying for or accessing health care. Please respond in at least four sentences."),
  list(phase = "engagement_question", text = "When you hear publicly funded health insurance, what institutional model do you imagine, and what outcome would make you call it successful? Please answer with at least four sentences."),
  list(phase = "exploration_question", text = "Should coverage be universal for all residents, or targeted to specific groups? Explain your principles and practical concerns in detail, with at least four sentences."),
  list(phase = "probing_focused", text = "What financing mix would you accept: payroll tax, progressive income tax, wealth tax, consumption tax, or blended mechanisms? Explain why in at least four sentences."),
  list(phase = "exploration_question", text = "Which benefits must be guaranteed: primary care, emergency, mental health, maternal care, pediatric care, dental, vision, prescription drugs, and long term care? Please prioritize with reasons."),
  list(phase = "exploration_question", text = "How much role should private insurance keep if a public system expands? Compare single-payer, public option, and tightly regulated multi-payer designs."),
  list(phase = "probing_focused", text = "Discuss the strongest argument against your own preferred model. What downside worries you most, and how would you mitigate it?"),
  list(phase = "exploration_question", text = "How should policy address equity gaps across class, race, rural geography, disability, and immigration status while preserving administrative clarity?"),
  list(phase = "exploration_question", text = "Who should govern the system, federal or state institutions, and what anti-corruption and anti-waste controls are indispensable?"),
  list(phase = "engagement_question", text = "Assume this reform is politically contested. What message would persuade skeptical voters without misrepresenting trade-offs?"),
  list(phase = "ending_question", text = "State your final package in three elements: financing, coverage architecture, and implementation safeguards. Defend feasibility and legitimacy."),
  list(phase = "closing")
)

prompts <- FocusGroup::get_default_prompt_templates()
prompts$participant_utterance_subtle_persona <- paste0(
  prompts$participant_utterance_subtle_persona,
  "\n8. Length requirement: write 4 to 8 sentences and approximately 90 to 180 words.",
  "\n9. Include at least one concrete policy mechanism and one explicit trade-off.",
  "\n10. If you agree with another participant, add one original caveat or extension so your response is not redundant.",
  "\n11. Avoid one-sentence or slogan-style responses."
)

prompts$moderator_exploration_question <- paste0(
  prompts$moderator_exploration_question,
  "\nAdditional moderator instruction: if answers become short, explicitly ask for detail, examples, and trade-off reasoning."
)

prompts$moderator_engagement_question <- paste0(
  prompts$moderator_engagement_question,
  "\nAdditional moderator instruction: request multi-sentence responses with policy specificity."
)
```

## Run Live Simulation (No Fabricated Fallback)

```{r run-live}
flow_mode <- tolower(Sys.getenv("FG_DEMO_FLOW", "round_robin"))
if (!(flow_mode %in% c("round_robin", "probabilistic", "desire_based"))) {
  flow_mode <- "round_robin"
}

fg <- FocusGroup$new(
  topic = fg_topic,
  purpose = fg_purpose,
  agents = agents_named,
  moderator_id = moderator_id,
  turn_taking_flow = create_conversation_flow(
    flow_mode,
    agents_named,
    moderator_id,
    flow_params = list(min_desire_threshold = 2)
  ),
  question_script = script,
  prompt_templates = prompts,
  llm_config_admin = cfg_admin,
  max_tokens_config = list(utterance = 380, moderator = 500, desire = 22)
)

run_result <- tryCatch(
  {
    num_turns_live <- suppressWarnings(as.integer(Sys.getenv("FG_DEMO_TURNS", "24")))
    if (is.na(num_turns_live) || num_turns_live < 20) num_turns_live <- 24L
    fg$run_simulation(num_turns = num_turns_live, verbose = TRUE)
  },
  error = function(e) {
    stop("Live API simulation failed: ", conditionMessage(e))
  }
)

if (length(fg$conversation_log) == 0) {
  stop("Simulation returned an empty conversation log. No fallback is used in this file.")
}

turn_df <- build_turn_dataframe(fg$conversation_log)
analysis_core <- fg$analyze()
full_transcript <- analysis_core$full_transcript
```

## Participant Composition Check

```{r composition}
kable(participant_meta, caption = "Participant metadata used for subgroup interpretation")

party_counts <- participant_meta |>
  count(party_group, name = "n_participants") |>
  arrange(desc(n_participants))

education_counts <- participant_meta |>
  count(education, name = "n_participants") |>
  arrange(desc(n_participants))

kable(party_counts, caption = "Party composition")
kable(education_counts, caption = "Education composition")
```

## Expandable Transcript

```{r transcript, results='asis'}
cat("### Full Transcript\n\n")

for (i in seq_len(length(fg$conversation_log))) {
  entry <- fg$conversation_log[[i]]
  speaker <- entry$speaker_id %||% "Unknown"
  phase <- entry$phase %||% "unknown"
  stamp <- format(as.POSIXct(entry$timestamp %||% Sys.time()), "%H:%M:%S")
  body_class <- if (isTRUE(entry$is_moderator) || identical(speaker, moderator_id)) "moderator" else "participant"
  speaker_type <- if (body_class == "moderator") "Moderator" else "Participant"
  body_text <- escape_html(entry$text %||% "")
  body_text <- gsub("\n", "<br/>", body_text, fixed = TRUE)

  cat('<details class="turn-card">\n')
  cat(
    '<summary>',
    "<strong>Turn ", i, " | ", speaker, "</strong>",
    ' <span class="turn-meta">(',
    speaker_type, " | ", tools::toTitleCase(phase), " | ", stamp,
    ')</span></summary>\n',
    sep = ""
  )
  cat('<div class="turn-body ', body_class, '">', body_text, "</div>\n", sep = "")
  cat("</details>\n\n")
}
```

## Conversation Diagnostics

```{r diagnostics}
speaker_stats <- analysis_core$speaker_stats
kable(speaker_stats, caption = "Speaker-level participation statistics")

phase_stats <- turn_df |>
  count(phase, name = "turns") |>
  mutate(share = round(100 * turns / sum(turns), 1)) |>
  arrange(desc(turns))
kable(phase_stats, caption = "Turn distribution by phase")

avg_turn_length <- turn_df |>
  filter(!is_moderator) |>
  summarise(
    avg_words = round(mean(word_count), 1),
    median_words = round(median(word_count), 1),
    max_words = max(word_count)
  )
kable(avg_turn_length, caption = "Participant turn length diagnostics")
```

## Education and Party Signal Tables

```{r subgroup-signals}
turn_with_meta <- turn_df |>
  filter(!is_moderator) |>
  left_join(participant_meta, by = c("speaker_id" = "participant_id"))

support_pat <- regex("universal|public option|single payer|coverage guarantee|health care as a right|expand public insurance", ignore_case = TRUE)
concern_pat <- regex("tax burden|tax increase|wait time|bureaucracy|government inefficien|ration|provider shortage|fiscal risk", ignore_case = TRUE)
market_pat <- regex("private insurance|competition|consumer choice|market discipline|employer plan", ignore_case = TRUE)

signal_by_party <- turn_with_meta |>
  mutate(
    support_hits = str_count(text, support_pat),
    concern_hits = str_count(text, concern_pat),
    market_hits = str_count(text, market_pat),
    net_support = support_hits - concern_hits
  ) |>
  group_by(party_group) |>
  summarise(
    n_turns = n(),
    avg_words = round(mean(word_count), 1),
    support_hits = sum(support_hits),
    concern_hits = sum(concern_hits),
    market_hits = sum(market_hits),
    net_support = sum(net_support),
    .groups = "drop"
  ) |>
  arrange(desc(net_support))

signal_by_education <- turn_with_meta |>
  mutate(
    support_hits = str_count(text, support_pat),
    concern_hits = str_count(text, concern_pat),
    market_hits = str_count(text, market_pat),
    net_support = support_hits - concern_hits
  ) |>
  group_by(education) |>
  summarise(
    n_turns = n(),
    avg_words = round(mean(word_count), 1),
    support_hits = sum(support_hits),
    concern_hits = sum(concern_hits),
    market_hits = sum(market_hits),
    net_support = sum(net_support),
    .groups = "drop"
  ) |>
  arrange(desc(net_support))

kable(signal_by_party, caption = "Lexical policy orientation by party group")
kable(signal_by_education, caption = "Lexical policy orientation by education group")
```

## Standard FocusGroup Analyses

```{r package-analyses}
topic_res <- fg$analyze_topics(num_topics = 5, top_n_terms = 10)
tfidf_res <- fg$analyze_tfidf(top_n_terms = 8)
readability_res <- fg$analyze_readability(measures = c("Flesch", "Flesch.Kincaid", "ARI"))
theme_res <- fg$analyze_themes(llm_config = cfg_admin)

print(topic_res)
kable(tfidf_res, caption = "Top TF-IDF terms by speaker")
kable(readability_res, caption = "Readability scores by speaker")

cat("### Theme analysis from package method\n\n")
cat(as.character(theme_res))
```

## Final DeepSeek Comparative Call with Personas and Transcript

```{r final-deepseek-comparison}
persona_blocks <- vapply(participant_ids, function(pid) {
  row <- participant_meta |> filter(participant_id == pid)
  party <- row$party_group[1] %||% "Unknown"
  edu <- row$education[1] %||% "Unknown"
  ideo <- row$ideology[1] %||% "Unknown"
  persona <- fg$agents[[pid]]$persona_description %||% ""
  persona <- substr(persona, 1, 2500)
  paste0(
    "Participant: ", pid, "\n",
    "Party group: ", party, "\n",
    "Education: ", edu, "\n",
    "Ideology: ", ideo, "\n",
    "Persona description:\n", persona, "\n"
  )
}, character(1))

persona_text <- paste(persona_blocks, collapse = "\n--------------------\n")

analysis_prompt <- paste0(
  "You are a political behavior analyst. You are given a focus group transcript and persona data.\n",
  "Task: extract subgroup differences with explicit evidence.\n\n",
  "Required sections:\n",
  "1) Party comparison: Democrats vs Republicans vs Independents on publicly funded health insurance.\n",
  "2) Education comparison: lower vs higher education patterns and whether sophistication changes policy design preferences.\n",
  "3) Areas of consensus and conflict, including disagreements about taxes, private insurance role, wait times, and governance trust.\n",
  "4) Detect if support appears artificially uniform; if yes, diagnose why and specify which participants still expressed caution.\n",
  "5) Evidence table: participant ID, subgroup labels, position summary, and one quote fragment or close paraphrase.\n",
  "6) Method note: explain limits, including sample size and simulated setting.\n\n",
  "Write analytically and avoid generic language.\n\n",
  "<personas>\n", persona_text, "\n</personas>\n\n",
  "<transcript>\n", full_transcript, "\n</transcript>\n"
)

cfg <- LLMR::llm_config(
  provider = "deepseek",
  model = "deepseek-v4-pro",
  thinking = "enabled",
  reasoning_effort = "high"
)

OUT <- LLMR::call_llm(
  config = cfg,
  messages = list(list(role = "user", content = analysis_prompt))
)

final_text <- OUT$text %||% unwrap_llm_text(OUT)

cat("### Direct DeepSeek synthesis over personas + full transcript\n\n")
cat(final_text)
```

## Closing Notes

- This file intentionally requires live API execution.
- If your provider key is configured, the output should contain both simulated dialogue and LLM-based subgroup interpretation.
- The final section is the one to cite in presentation when discussing party and education heterogeneity in policy positions.
