---
title: "FocusGroup Package Demo"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

## Setup

```{r setup}
library(FocusGroup)

config <- LLMR::llm_config(
  provider = "deepseek",
  model = "deepseek-chat"
)

topic <- "The future of remote work: Balancing productivity and well-being"
```

## Run the Session

`run_focus_group()` requires an explicit configuration. The `guide` can contain
phase counts or the full moderator script.

```{r simulation}
result <- run_focus_group(
  topic = topic,
  n_participants = 3,
  guide = list(
    Opening = "Welcome the group and state the ground rules.",
    Icebreaker = "What has your experience with remote work been?",
    Engagement = c(
      "How has remote work affected your productivity?",
      "What benefits and costs have you experienced?"
    ),
    Exploration = c(
      "How do you maintain boundaries between work and home?",
      "Which policy or practice would improve remote work?"
    ),
    Closing = "Thank the participants and close the session."
  ),
  flow = "desire_based",
  config = config,
  seed = 110,
  message_mode = "roleflip"
)

print(result)
```

## Transcript and Usage

```{r transcript}
head(result$transcript)
result$participants
result$usage
```

The transcript uses `message_id` for message order and `round` for the
moderator cycle. Messages responding to one moderator question share a round.

## Analysis

Descriptive analysis runs offline. Passing `config` opts in to thematic model
analysis.

```{r analysis}
descriptive <- analyze_focus_group(result, include_plots = FALSE)
print(descriptive)

model_analysis <- analyze_focus_group(
  result,
  config = config,
  include_plots = FALSE
)
cat(model_analysis$themes)
```

## Interpretation

The transcript records model behavior under this guide, persona construction,
flow, and configuration. It is not an estimate of public opinion.
